Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-11-26 doc/Paper/main.tex
Addressed comments from Eric.

\abstract{We analyze the clustering of galaxies in the first public data release of the Hyper Suprime-Cam Subaru Strategic Program. Despite the relatively small footprints of the observed fields, the data are an excellent proxy for the very deep photometric datasets as will be acquired by the Large Synoptic Survey Telescope, and are therefore an ideal test bed for the analysis methods being implemented by the Dark Energy Science Collaboration. We select a magnitude limited sample with $i<24.5$ and analyze it in four tomographic redshift bins covering the range $0.15\lesssim z\lesssim1.5$. We carry out a Fourier-space analysis of the two-point clustering of this sample, including all auto- and cross-correlations between bins. We demonstrate the use of map-level deprojection methods to account for non-physical fluctuations in the galaxy number density caused by observational systematics. Through a halo occupation distribution analysis, we place constraints on the characteristic halo masses of this sample as a function of redshift, finding a good fit up to scales $k_{\rm max}=1\,{\rm Mpc}^{-1}$, including both auto- and cross-correlations. Our results show monotonically decreasing average halo masses, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a flux-limited sample, consistent with previous analyses. In terms of photometric redshift systematics, we show that additional care is needed in order to marginalize over uncertainties in the redshift distribution in galaxy clustering, even for samples of this small size, and that these uncertainties can be significantly constrained by including cross-bin correlations. We are able to make a $\sim3\sigma$ detection of the effects of lensing magnification in the HSC data. Our results are stable to variations in the amplitude of density fluctuations $\sigma_8$ and the cold dark matter abundance $\Omega_c$ and we find constraints that agree well with measurements from Planck and low-redshift probes. Finally, we use our analysis pipeline to study the clustering of galaxies as a function of limiting flux, and provide a simple fitting function for the linear galaxy bias for magnitude limited samples as a function of limiting magnitude and redshift.}

@@

      In Fig.~\ref{fig:HOD-redshift}, we additionally show the large-scale galaxy bias as a function of redshift derived from our best-fit HOD model. As can be seen, the large-scale bias scales approximately as $b(z) \propto \sfrac{1}{D(z)}$, where $D(z)$ denotes the linear growth factor. This behavior is consistent with e.g. studies of galaxies in the DEEP2 survey, which found constant clustering strengths as a function of redshift (and thus $b(z) \propto \sfrac{1}{D(z)}$) for magnitude-limited samples \cite{Coil:2004}.

@@

      \begin{figure}
        \begin{center}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/M_halo_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/bias-redshift_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/bias-growth_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \caption{Redshift dependence of the mean halo mass $\bar{M}_{\mathrm{halo}}$ and the large-scale galaxy bias $b_{\mathrm{HOD}}$ as derived from our best-fit HOD model. The third panel illustrates the empirical scaling $b(z) \propto \sfrac{1}{D(z)}$ derived from the galaxy sample used in this analysis.}
          \label{fig:HOD-redshift}
        \end{center}
      \end{figure}

@@

    To this end, we split our fiducial galaxy sample into four brighter subsamples with limiting magnitudes $i_{\mathrm{corr}} < 20.5$, $i_{\mathrm{corr}} < 21.5$, $i_{\mathrm{corr}} < 22.5$ and $i_{\mathrm{corr}} < 23.5$, respectively. We then subdivide each of these samples into four redshift bins of approximately equal galaxy number and use our pipeline to compute constraints on HOD parameters for our fiducial analysis variant. We find all subsamples to be well-fit by our fiducial model described in Sec.~\ref{ssec:methods.theory}. In fact, the reduced $\chi^{2}$s for some subsamples are quite low and thus lead to correspondingly high $p$-values, which seem to indicate the presence of overfitting and/or covariance overestimation in our brighter magnitude samples. However, for all subsamples, the distribution of the fit residuals is consistent with a Gaussian and we do not find any significant evidence that our covariance is overestimated by comparing the consistency of power spectra across different HSC fields. We therefore use the results of these HOD fits to derive constraints on the large-scale galaxy bias as a function of redshift for each sample. We determine the bias from the mean of the posterior distribution, as we find it to yield more stable constraints than the best-fit value. The results for the bias evaluated at the effective redshift of each tomographic bin are shown in the upper left panel of Fig.~\ref{fig:magnitude-cuts-bias}, alongside the results for our fiducial sample. As can be seen, we find the linear bias to increase for decreasing limiting magnitude, as expected. Furthermore, the redshift-dependence of the bias can be roughly approximated as inversely proportional to the growth factor, $b(z) \propto \sfrac{1}{D(z)}$, for all samples, as illustrated in the upper right panel of Fig.~\ref{fig:magnitude-cuts-bias}. Motivated by these results, we make the following Ansatz to derive a fitting function for the large-scale galaxy bias as a function of redshift and limiting magnitude:
    \begin{equation}
      b(z, m_{\mathrm{lim}}) = \bar{b}(m_{\mathrm{lim}}) D(z)^{\alpha},
      \label{eq:bias-fit-func}
    \end{equation}
    where we have assumed the bias to be separable in $z$ and $m_{\mathrm{lim}}$, and the parameter $\alpha$ accounts for deviations from the simple inverse growth function proportionality. We determine the values of $\alpha$ and $\bar{b}(m_{\mathrm{lim}})$ in a multi-step weighted, linear least squares fit: we first fit Eq.~\ref{eq:bias-fit-func} to all five samples separately and determine our fiducial $\alpha$ as the median of these values\footnote{We note that we find slightly better results using the median, which is the reason for preferring it over the mean.}. We then fit Eq.~\ref{eq:bias-fit-func} to all samples again, obtaining constraints on $\bar{b}(m_{\mathrm{lim}})$ for $\alpha$ fixed to our fiducial value. In a last step, we fit a linear function to $\bar{b}(m_{\mathrm{lim}})$, finally yielding
    \begin{equation}
      \begin{aligned}
        \alpha &= -1.30 \pm 0.19\footnotemark,\\
        \bar{b}(m_{\mathrm{lim}}) &= b_{1}(m_{\mathrm{lim}}-24) + b_{0},
        \label{eq:bias-fit-func-result}
      \end{aligned}
    \end{equation}
    \footnotetext{We note that we used $\sigma(\tilde{x}) = 1.253 \; \sigma(\bar{x})$, where $\tilde{x}$ denotes the median and $\bar{x}$ is the mean of a given sample.}
    where $b_{1} = -0.0624 \pm 0.0070$ and $b_{0} = 0.8346 \pm 0.161$. The results from this fitting function are shown as solid lines in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the values obtained from the HOD fit.

>>>>

\abstract{We analyze the clustering of galaxies in the first public data release of the Hyper Suprime-Cam Subaru Strategic Program. Despite the relatively small footprints of the observed fields, the data are an excellent proxy for the very deep photometric datasets as will be acquired by the Large Synoptic Survey Telescope, and are therefore an ideal test bed for the analysis methods being implemented by the Dark Energy Science Collaboration. We select a magnitude limited sample with $i<24.5$ and analyze it in four tomographic redshift bins covering the range $0.15\lesssim z\lesssim1.5$. We carry out a Fourier-space analysis of the two-point clustering of this sample, including all auto- and cross-correlations between bins. We demonstrate the use of map-level deprojection methods to account for non-physical fluctuations in the galaxy number density caused by observational systematics. Through a halo occupation distribution analysis, we place constraints on the characteristic halo masses of this sample as a function of redshift, finding a good fit up to scales $k_{\rm max}=1\,{\rm Mpc}^{-1}$, including both auto- and cross-correlations. Our results show monotonically decreasing average halo masses with increasing redshift, which can be interpreted in terms of the drop-out of red galaxies at high redshifts for a flux-limited sample, consistent with previous analyses. In terms of photometric redshift systematics, we show that additional care is needed in order to marginalize over uncertainties in the redshift distribution in galaxy clustering, even for samples of this small size, and that these uncertainties can be significantly constrained by including cross-bin correlations. We are able to make a $\sim3\sigma$ detection of the effects of lensing magnification in the HSC data. Our results are stable to variations in the amplitude of density fluctuations $\sigma_8$ and the cold dark matter abundance $\Omega_c$ and we find constraints that agree well with measurements from Planck and low-redshift probes. Finally, we use our analysis pipeline to study the clustering of galaxies as a function of limiting flux, and provide a simple fitting function for the linear galaxy bias for magnitude limited samples as a function of limiting magnitude and redshift.}

@@

      In Fig.~\ref{fig:HOD-redshift}, we additionally show the large-scale galaxy bias as a function of redshift derived from our best-fit HOD model. As can be seen, the large-scale bias scales approximately as $b(z) \propto \sfrac{1}{D(z)}$ (to within approximately $5 \%$), where $D(z)$ denotes the linear growth factor. This behavior is consistent with e.g. studies of galaxies in the DEEP2 survey, which found constant clustering strengths as a function of redshift (and thus $b(z) \propto \sfrac{1}{D(z)}$) for magnitude-limited samples \cite{Coil:2004}.

@@

      \begin{figure}
        \begin{center}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/M_halo_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/bias-redshift_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \begin{subfigure}{0.32\textwidth}
          \includegraphics[width=\textwidth]{figures/bias-growth_mPk=HOD_fix=alpha-fc-sigmaM_HOD=zevol_fit=pz-shifts-pz-widths+prior=0p2_fit=auto+cross_cosmo=const_cov=G+NG+SSC-LINBIAS_HOD-param=zfid_clfit=HOD-zevol.pdf}
          \end{subfigure}
          \caption{Redshift dependence of the mean halo mass $\bar{M}_{\mathrm{halo}}$ and the large-scale galaxy bias $b_{\mathrm{HOD}}$ as derived from our best-fit HOD model. The third panel illustrates the empirical scaling $b(z) \propto \sfrac{1}{D(z)}$, which is consistent with the bias-redshift relation derived from the galaxy sample used in this analysis to within approximately $5 \%$.}
          \label{fig:HOD-redshift}
        \end{center}
      \end{figure}

@@

    To this end, we split our fiducial galaxy sample into four brighter subsamples with limiting magnitudes $i_{\mathrm{corr}} < 20.5$, $i_{\mathrm{corr}} < 21.5$, $i_{\mathrm{corr}} < 22.5$ and $i_{\mathrm{corr}} < 23.5$, respectively. We then subdivide each of these samples into four redshift bins of approximately equal galaxy number and use our pipeline to compute constraints on HOD parameters for our fiducial analysis variant. We find all subsamples to be well-fit by our fiducial model described in Sec.~\ref{ssec:methods.theory}. In fact, the reduced $\chi^{2}$s for some subsamples are quite low and thus lead to correspondingly high $p$-values, which seem to indicate the presence of overfitting and/or covariance overestimation in our brighter magnitude samples. However, for all subsamples, the distribution of the fit residuals is consistent with a Gaussian and we do not find any significant evidence that our covariance is overestimated by comparing the consistency of power spectra across different HSC fields. We therefore use the results of these HOD fits to derive constraints on the large-scale galaxy bias as a function of redshift for each sample. We determine the bias from the mean of the posterior distribution, as we find it to yield more stable constraints than the best-fit value. The results for the bias evaluated at the effective redshift of each tomographic bin are shown in the upper left panel of Fig.~\ref{fig:magnitude-cuts-bias}, alongside the results for our fiducial sample. As can be seen, we find the linear bias to increase for decreasing limiting magnitude, as expected. Furthermore, the redshift-dependence of the bias can be roughly approximated as inversely proportional to the growth factor, $b(z) \propto \sfrac{1}{D(z)}$, for all samples, as illustrated in the upper right panel of Fig.~\ref{fig:magnitude-cuts-bias}. Motivated by these results, we make the following Ansatz to derive a fitting function for the large-scale galaxy bias as a function of redshift and limiting magnitude:
    \begin{equation}
      b(z, m_{\mathrm{lim}}) = \bar{b}(m_{\mathrm{lim}}) D(z)^{\alpha},
      \label{eq:bias-fit-func}
    \end{equation}
    where we have assumed the bias to be separable in $z$ and $m_{\mathrm{lim}}$, and the parameter $\alpha$ accounts for deviations from the simple inverse growth function proportionality. We determine the values of $\alpha$ and $\bar{b}(m_{\mathrm{lim}})$ in a multi-step weighted, linear least squares fit: we first fit Eq.~\ref{eq:bias-fit-func} to all five samples separately and determine our fiducial $\alpha$ as the median of these values\footnote{We note that we find slightly better results using the median, which is the reason for preferring it over the mean.}. We then fit Eq.~\ref{eq:bias-fit-func} to all samples again, obtaining constraints on $\bar{b}(m_{\mathrm{lim}})$ for $\alpha$ fixed to our fiducial value. In a last step, we fit a linear function to $\bar{b}(m_{\mathrm{lim}})$, finally yielding
    \begin{equation}
      \begin{aligned}
        \alpha &= -1.30 \pm 0.19\footnotemark,\\
        \bar{b}(m_{\mathrm{lim}}) &= b_{1}(m_{\mathrm{lim}}-24) + b_{0},
        \label{eq:bias-fit-func-result}
      \end{aligned}
    \end{equation}
    \footnotetext{We note that we use $\sigma(\tilde{x}) = 1.253 \; \sigma(\bar{x})$, where $\tilde{x}$ denotes the median and $\bar{x}$ is the mean of a given sample.}
    where $b_{1} = -0.0624 \pm 0.0070$ and $b_{0} = 0.8346 \pm 0.161$. The results from this fitting function are shown as solid lines in Fig.~\ref{fig:magnitude-cuts-bias}, alongside the values obtained from the HOD fit.
